---
title: "Changing parameters in the model of unguarded X effects"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

```{r setup}

library(ggplot2)
library(dplyr)


# Set up custom theme for ggplots
custom_theme <-
  list(
    theme_bw() +
      theme(
        panel.border = element_blank(),
        axis.line = element_line(),
        text = element_text(size = 25),
        legend.position = "bottom",
        strip.background = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
      )
  )

```

### Starting assumptions – Effects under sex-equal mutation and selection: 


 $s_{f,i} = s_{m,i}$  and $\mu_{f,i} = \mu{m,i}$, with $h_{i}$, $\mu_{j,i}$, and $\alpha_i$ independently distributed. 
 
 
$$\frac{{\bar{W}}_{f,X}}{{\bar{W}}_{m,X}}\approx\exp{\left(\sum_{i=1}^{n_X}{\left(s_{m,i}\alpha_{m,i}-2h_is_{f,i}\alpha_{f,i}\right)\frac{2\mu_{f,i}+\mu_{m,i}}{2h_is_{f,i}+s_{m,i}}}\right)}$$


```{r first_eq}

sex_bias_survive_ratio = function(s_m,
                                  s_f,
                                  alpha_m,
                                  alpha_f,
                                  mu_f,
                                  mu_m,
                                  h){
  
return(
  
exp(
    sum(
         (s_m*alpha_m - 2*h*s_f*alpha_f)*(2*mu_f+mu_m)/(2*h*s_f+s_m)
                                                                    )
                                                                      )
  
)
  
}


```

### Effects of sex-biased purifying selection



Assuming that mutations similarly affect viability in each sex ($\alpha_{f,i},s_{f,i}=\alpha_{m,i}s_{m,i}$), and  $$ \beta = \frac{s_{m,i}}{s_{f,i}} = \frac{\alpha_{f,i}}{\alpha_{m,i}} $$ is relatively constant among mutations affecting viability, we can generalize to cases involving sex-biased purifying selection (i.e., $\beta \neq 1$):

$$ \frac{{\bar{W}}_{f,X}}{{\bar{W}}_{m,X}}\approx\exp{\left(3U_X\bar{\alpha}\left(\frac{1-2\bar{h}}{1+2\bar{h}}+\frac{8var{\left(h\right)}}{\left(1+2\bar{h}\right)^3}\right)\right)} $$


```{r second_eq}


sex_bias_survive_ratio_2 = function(u_x,
                                    alpha_hat,
                                    h){
  
  h_hat = mean(h)
  
  w_ratio = exp(
    
                3*u_x*alpha_hat*(
                                (1-h_hat)/(1+h_hat) + 
                                                      (8*var(h))/((1+2*h_hat)^3) 
                                                                                 )
  
                                                                                   )  


return(
        data.frame(w_ratio = w_ratio,
                   alpha_hat = alpha_hat, 
                   var_h = var(h),
                   h_hat = h_hat,
                   u_x = u_x) 
)
  
}


sex_bias_survive_ratio_2(u_x = 0.1, 
                         alpha_hat = seq(from = 0, to = 1, by = 0.001), 
                         h = rep(1,100)) %>% 
  ggplot(aes(x=alpha_hat,y=w_ratio)) + 
  geom_point() + 
  custom_theme



sex_bias_survive_ratio_2(u_x = 0.1, 
                         alpha_hat = seq(from = 0, to = 1, by = 0.01), 
                         h = rep(0,100)) %>% 
  ggplot(aes(x=alpha_hat,y=w_ratio)) + 
  geom_point() + 
  custom_theme

```


$$ \frac{{\bar{W}}_{f,X}}{{\bar{W}}_{m,X}}\approx\exp{\left(3U_X\frac{{\bar{\alpha}}_f\left(1-2\bar{h}\right)}{\beta+2\bar{h}}\left(1+\frac{4var{\left(h\right)}\left(\beta+1\right)}{\left(1-2\bar{h}\right)\left(\beta+2\bar{h}\right)^2}\right)\right)} $$

```{r third_eq}

sex_bias_survive_ratio_3 = function(u_x,
                                    alpha_hat_f,
                                    h,
                                    beta){
  
  h_hat = mean(h)
  
  
return(
  exp(
  3*
    u_x*(
          (
            (alpha_hat_f)*(1-2*h_hat)
                                      )/(
                                          beta + 2*h_hat
                                                         )
                                                          )*
    (1 + 
         (
           (4*var(h))(beta+1)
                              )/(
                                   (1-2*h_hat)*(
                                                  (beta + 2*h_hat)^2)
                                                                      )
                                                                       )
  
  )
)
  
                                    }

```


