---
title: "Summary information on the male-biased mutation rate estimate dataset"
author: "Isobel Beasley & Yasmine McDonough"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

**Introduction** This workflowr page contains summary information / investigations on the dataset of male-biased estimates we compiled from published works. 

### Set up

```{r setup, message = FALSE, warning=FALSE}

library(ggplot2)
suppressMessages(library(dplyr))
library(reactable)
library(tidyr)

# read in sex bias mutation data.frame
sex_bias_mut_df = readxl::read_excel("data/Sex_bias_mutation_rate.xlsx", sheet = "Sex-bias estimates")

# what does the sex bias in mutation rate estimate look like: 
head(sex_bias_mut_df) %>% reactable()

# Set up custom theme for ggplots
custom_theme <-
  list(
    theme_bw() +
      theme(
        panel.border = element_blank(),
        axis.line = element_line(),
        text = element_text(size = 16),
        legend.position = "bottom",
        strip.background = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
      )
  )

```

# What data have we collected / collated?

## Number of unique studies included

```{r n_unique_studies}

sex_bias_mut_df %>% 
  select(authors,year) %>% 
  distinct() %>% 
  nrow()

```

## When where these studies published?

```{r when_studies_published, fig.height = 8, fig.width = 12}

sex_bias_mut_df %>%   
  select(authors,year) %>% 
  distinct() %>% 
  ggplot(aes(x=year)) + 
  geom_histogram() + 
  custom_theme  + 
  labs(x = "Year published*",
       y = "Number",
       title = "What year were our collated studies published?")


```

* Or preprinted, in the case of included preprints

## How many distinct estimates per year?

```{r n_samples_per_year, fig.height = 8, fig.width = 12}

sex_bias_mut_df %>%   
  select(authors,year,scientific_name) %>% 
  distinct() %>% 
  ggplot(aes(x=year)) + 
  geom_histogram() + 
  custom_theme  + 
  labs(x = "Year published*",
       y = "Number",
       title = "How many unique estimtates (study, and organism specific) per year?")


```

* Or preprinted, in the case of included preprints

## Number of estimates of sex-biased mutation rate per taxonomic class

```{r n_by_taxo_class}

sex_bias_mut_df %>% 
  group_by(taxonomic_class) %>% 
  summarise(n = n()) %>% 
  reactable()

```

## Number of estimates per method used to calculate the sex-biased ratio in mutation rate

```{r n_by_method}

sex_bias_mut_df %>% 
  group_by(method_used_to_calculate_alpha) %>% 
  summarise(n = n()) %>% 
  reactable(defaultPageSize = 50)

```

## Number of estimates per loci type used to calculate the sex-biased ratio in mutation rate

```{r n_by_loci_type}

sex_bias_mut_df %>% 
  group_by(loci_type_for_estimate) %>% 
  summarise(n = n()) %>% 
  reactable(defaultPageSize = 50)

# Number of unique studies

sex_bias_mut_df %>% 
  group_by(authors,year) %>% 
  sample_n(1) %>% 
  ungroup() %>% 
  group_by(taxonomic_class,loci_type_for_estimate) %>% 
  summarise(n = n()) %>% 
  reactable(defaultPageSize = 50)

```

## Number of estimates by both taxonomic class and method used to calculate

```{r n_by_method_and_taxo_class}

sex_bias_mut_df %>% 
  group_by(taxonomic_class,method_used_to_calculate_alpha) %>% 
  summarise(n = n()) %>% 
  reactable(defaultPageSize = 50)

```

# Distribution of sex-biased mutation estimates

For the final figure 1 (and the code used to produce this figure) - see this [workflowr page](https://ijbeasley.github.io/unguarded_x_journal_club/figure_1.html).

## Across all taxonomic classes

```{r dist_sex_est, fig.height = 8, fig.width = 12}

sex_bias_mut_df %>% 
  ggplot(aes(x=alpha_point_estimate)) + 
  geom_histogram(bins = 50) + 
  labs(title = "Distribution of estimated sex-biased mutation rate",
        x = "Point estimate of the male-biased mutation rate (alpha)",
       y= "Count") + 
  custom_theme + 
  geom_vline(xintercept = 1, col = "red", linetype = 2, size = 2) + 
  annotate(
  "text",
  x = 3,
  y = 35,
  size = 9,
  hjust = 0,
  col = "red", 
  label = "Line of equal mutation rate across the sexes"
)

summary(sex_bias_mut_df$alpha_point_estimate)

sex_bias_mut_df %>% 
  ggplot(aes(x=alpha_point_estimate)) +
  geom_boxplot() + 
  custom_theme+ 
   labs(title = "Distribution of estimated sex-biased mutation rate",
        x = "Point estimate of the male-biased mutation rate (alpha)") 

```

### After sampling down multiple estimates on the same sample (i.e. same study, and organism)

```{r sampling_down_est_dist_est_in_mut_bias, fig.height = 8, fig.width = 12}

sex_bias_mut_df %>% 
  group_by(authors,year,scientific_name) %>% 
  sample_n(1) %>% 
  ggplot(aes(x=alpha_point_estimate)) + 
  geom_histogram(bins = 50) + 
  labs(title = "Distribution of estimated sex-biased mutation rate",
        x = "Point estimate of the male-biased mutation rate (alpha)",
       y= "Number of estimates") + 
  custom_theme + 
  geom_vline(xintercept = 1, col = "red", linetype = 2, size = 2) + 
  annotate(
  "text",
  x = 3,
  y = 35,
  size = 9,
  hjust = 0,
  col = "red", 
  label = "Line of equal mutation rate across the sexes"
)


```

## By Taxonomic Class

```{r sex_bias_by_taxo_class, fig.height = 8, fig.width = 12}

df= sex_bias_mut_df %>% 
  filter(alpha_point_estimate<20) %>% 
  ggplot(aes(x=alpha_point_estimate)) + 
  geom_histogram(bins = 25) + 
  labs(title = "Distribution of sex-biased mutation rate estimtates",
        x = "Point estimate of the male-biased mutation rate (alpha)",
       y= "Number of estimates") + 
  custom_theme + 
  geom_vline(xintercept = 1, col = "red", linetype = 2, size = 2) + 
  facet_wrap(~taxonomic_class)

n_by_taxa = sex_bias_mut_df %>% 
            group_by(taxonomic_class) %>% 
            summarise(n = n()) 

library(gridExtra)

df_2 = sex_bias_mut_df %>%   
  select(authors,year,scientific_name) %>% 
  distinct() %>% 
  ggplot(aes(x=year)) + 
  geom_histogram() + 
  custom_theme  + 
  labs(x = "Year published*",
       y = "Number of estimates",
       title = "How many unique estimates per year?")


df_3 = sex_bias_mut_df %>%   
  select(authors,year) %>% 
  distinct() %>% 
  ggplot(aes(x=year)) + 
  geom_histogram() + 
  custom_theme  + 
  labs(x = "Year published*",
       y = "Number of studies",
       title = "What year were our collated studies published?")


library(patchwork)

(df) / (df_2 + df_3)

```

## By method used to calculate estimate of male-biased mutation rate

```{r sex_bias_by_method, fig.height=20, fig.width = 30}

sex_bias_mut_df %>% 
  ggplot(aes(x=alpha_point_estimate)) + 
  geom_histogram() + 
  custom_theme + 
  facet_wrap(~method_used_to_calculate_alpha) + 
  theme(text = element_text(size = 25)) +   
  labs(title = "Distribution of estimated sex-biased mutation rate",
        x = "Point estimate of the male-biased mutation rate (alpha)",
       y= "Count")

```

## Confidence intervals - 

```{r}

df= sex_bias_mut_df %>% 
  filter(upper_lower_meaning == "Lower and upper estimates are the 95% confidence interval for alpha"|
         upper_lower_meaning == "Lower and upper estimates are 2 standard errors from the mean (rough 95% confidence interval)") %>% 
  mutate(se = (alpha_estimate_upper - alpha_point_estimate)/2) 


df %>% 
  select(taxonomic_class,se) %>% 
  mutate(row = row_number()) %>% 
  pivot_wider(names_from = taxonomic_class,
              values_from = se,
              values_fn = list)

```



